\documentclass[11pt,a4paper]{article}

% ---- packages (all standard TeX Live; compiles with pdflatex) ----
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[margin=2.4cm]{geometry}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{xcolor}
\usepackage{enumitem}
\usepackage{titlesec}

% colours must be defined BEFORE hyperref reads them in its options
\definecolor{NavyBlue}{RGB}{32,64,128}
\definecolor{Soft}{RGB}{245,247,250}

\usepackage[colorlinks=true,linkcolor=NavyBlue!70!black,urlcolor=NavyBlue!70!black,citecolor=NavyBlue!70!black]{hyperref}

% inline code that survives underscores etc.
\newcommand{\code}[1]{\texttt{#1}}
\newcommand{\boundary}[1]{\textcolor{NavyBlue!80!black}{\textbf{#1}}}

\titleformat{\section}{\Large\bfseries\color{NavyBlue!75!black}}{\thesection}{0.6em}{}
\titleformat{\subsection}{\large\bfseries\color{NavyBlue!60!black}}{\thesubsection}{0.6em}{}

\setlist[itemize]{leftmargin=1.3em,itemsep=2pt,topsep=2pt}

\newcolumntype{L}[1]{>{\raggedright\arraybackslash}p{#1}}

\title{\vspace{-1.5cm}\textbf{Vantage --- Security Architecture \&\\ Control Reference}}
\author{InfoSec / AppSec --- reference build\\\small Bajaj Life Insurance (IRDAI-regulated; ISO/IEC 27001:2022 context)}
\date{\today}

\begin{document}
\maketitle
\vspace{-0.8cm}

\noindent\fbox{\parbox{\dimexpr\textwidth-2\fboxsep-2\fboxrule}{%
\small \boundary{Hard product boundary.} Vantage \emph{scans and reports only}.
It detects, triages, and reports vulnerabilities; it does \textbf{not} exploit,
perform lateral movement, or auto-remediate. Every control in this document is
read as defense-in-depth around that boundary.}}

\section{Purpose and scope}
This document is the control reference for \emph{Vantage}, an AI-augmented
vulnerability console for a regulated life insurer. It enumerates each security
control: its \textbf{mechanism}, its \textbf{location} in the codebase, and how
it is \textbf{verified} (unit test / CI gate). It is intended as ISO/IEC
27001:2022 control evidence and as an engineering reference. The controls span
four enforcement surfaces --- the \textbf{database}, the \textbf{API}, the
\textbf{console UI}, and the \textbf{agent} (the Claude Code plugin).

\section{The hard boundary: scan-and-report only}
The boundary is enforced \emph{structurally}, not merely documented:
\begin{itemize}
  \item The pipeline \code{Phase} enum (\code{shared.py}) has \textbf{no}
        \code{exploit}/\code{lateral}/\code{remediate} state; the terminal phase
        is \code{report}.
  \item No scanner adapter exposes an offensive verb; the only methods are
        \code{preflight}, \code{launch}, \code{wait}, \code{fetch\_raw},
        \code{parse}. The live \code{launch}/\code{wait}/\code{fetch\_raw} are
        deliberately \code{NotImplementedError} in the reference build.
  \item This is asserted by an \textbf{executable invariant}
        (\code{orchestrator/test\_scope\_invariants.py}) and a CI grep guard ---
        a change that introduced an exploitation phase or verb fails the build.
\end{itemize}

\section{Control layers}

\subsection{Authorization to scan --- the scope gate}
A target is scanned only if it is in the HOD-approved inventory. The gate is
\textbf{fail-closed} and enforced at four independent points:
\begin{itemize}
  \item \boundary{Server} (\code{api/main.py}, \code{POST /api/scans}):
        \code{seed.asset\_by\_id(assetId)}; an unknown / non-approved asset is
        refused with HTTP \textbf{403 \code{out\_of\_scope}} \emph{before} any
        other validation, and the denial is audited
        (\code{SCAN\_DENIED\_OUT\_OF\_SCOPE}).
  \item \boundary{Orchestrator} (\code{activities.scope\_gate\_authorize}):
        resolves the target query against \code{assets WHERE
        approved\_in\_inventory}, mints an Ed25519-signed, time-boxed
        authorization token, and re-verifies it at the moment of use
        (\code{\_enforce\_scope}) as a TOCTOU defense; adapters additionally call
        \code{assert\_targets\_in\_scope}.
  \item \boundary{Console} (\code{StartScan.jsx}): the asset picker offers only
        approved assets, so an out-of-scope target cannot be selected.
  \item \boundary{Agent} (\code{plugin/hooks/scope\_guard.py}): a
        \code{PreToolUse} hook blocks scanner/recon shell commands or
        \code{request\_scan} aimed at a host not in the approved inventory.
\end{itemize}

\subsection{Authentication --- OIDC + signed session}
Implemented in \code{api/auth.py} against Active Directory / Entra via OIDC.
\begin{itemize}
  \item \textbf{Flow:} Authorization Code + \textbf{PKCE} (S256); \code{state} and
        \code{nonce} are carried in a signed, \code{httpOnly} transaction cookie,
        salt-namespaced from the session cookie so neither can be replayed as the
        other.
  \item \textbf{\code{id\_token} validation:} \code{JsonWebToken(["RS256"])} ---
        the algorithm allow-list is \textbf{pinned}, which structurally blocks
        the \code{alg:none} downgrade and HS/RS key-confusion attacks.
        \code{iss} and \code{aud} are essential and value-pinned, \code{exp} is
        essential (60\,s leeway), and \code{nonce} is matched. Keys are taken
        only from the IdP \textbf{JWKS} (via discovery).
  \item \textbf{Session:} a signed (\code{itsdangerous}) \code{httpOnly},
        \code{SameSite=Lax} cookie over \{\code{sub,name,email,roles,groups,exp}\};
        \emph{both} the cookie \code{max\_age} and the embedded \code{exp} are
        enforced; \code{Secure} in production.
  \item \textbf{Fail-closed:} with \code{AUTH\_REQUIRED=true} and no
        \code{SESSION\_SECRET}, signing \emph{raises} rather than using the known
        dev key --- so sessions are never forgeable in production.
  \item \textbf{Reference-build switch:} \code{AUTH\_REQUIRED} (default off) makes
        an unauthenticated caller a synthetic \code{dev} admin so the offline
        console and CI keep working; crucially, the \emph{actor is still derived
        server-side} in both modes.
\end{itemize}

\subsection{Authorization --- RBAC}
\begin{itemize}
  \item \code{require\_role(*roles)} dependency; \code{admin} is a wildcard;
        \textbf{401 (unauthenticated) always beats 403 (forbidden)} because the
        role check depends on \code{get\_current\_user}.
  \item Matrix: reads require any authenticated user; \code{analyst} for
        status / scan / exception / false-positive / retest; the \emph{exception's
        tier role} (CISO / RMC / Board) for approve--reject; \code{admin} for the
        escalation sweep.
  \item \textbf{Actor is server-derived} (\code{session\_actor(user)}) --- never
        read from the request body; this closes the spoofable-actor gap.
  \item \textbf{Report download is owner-scoped:}
        \code{entry["owner"] == user.sub} or \code{admin}, else 403.
  \item \textbf{State invariant:} \code{risk\_accepted} is reachable \emph{only}
        via an approved exception; \code{confirmed\_fp} only via the
        false-positive flow --- the generic status path cannot set either.
\end{itemize}

\subsection{Injection and input-validation defenses}
\begin{longtable}{L{3.1cm} L{11.2cm}}
\toprule
\textbf{Class} & \textbf{Defense} \\
\midrule
\endhead
XXE & All XML (Nmap / Nessus / Nikto) parsed with \code{defusedxml} (external
entities + billion-laughs blocked); the invariant test asserts no stdlib
\code{xml.etree} import anywhere. \\
\addlinespace
SQL injection & All values are parameterized (\code{\%s}); the only \code{\{table\}}
f-strings interpolate hardcoded internal literals, never request input. \\
\addlinespace
LDAP injection & The OIDC-claim email is shape-validated (rejects
\code{, \textbackslash\ * ( )} and NUL) \emph{and} \code{escape\_filter\_chars}
escaped (RFC 4515) before the \code{memberOf} filter; the search is skipped on
metacharacters. \\
\addlinespace
Open redirect & \code{\_safe\_next} requires a single-slash relative path; rejects
\code{//}, backslash, CR/LF/tab, and any non-empty \code{urlsplit().netloc}. \\
\addlinespace
Path traversal / SSRF & Report download maps \code{reportId}/\code{fmt} through an
in-memory registry to a server-built path (never the raw param); the MCP/hook
HTTP clients are pinned to a fixed \code{base\_url}, so adversarial paths stay
on-host. \\
\addlinespace
Code execution & JSON via \code{json.load} only --- no \code{eval} / \code{pickle}
/ \code{yaml.load}; the scope-guard hook \emph{inspects} shell strings, it never
executes them. \\
\bottomrule
\end{longtable}

\subsection{Cryptography and secrets}
\begin{itemize}
  \item \textbf{Report capability token:} \code{secrets.token\_urlsafe(24)}
        ($\approx$192 bits, unguessable), TTL-bounded ($\approx$1\,h), owner-bound,
        and not echoed into the download filename.
  \item \textbf{Dual-password PDF:} \code{pikepdf} AES-256 (R6) --- an \emph{open}
        password to view plus an \emph{owner} password that lifts the
        copy/modify/print restriction.
  \item \textbf{Signing keys:} Ed25519 for scope-authorization tokens (vault-held);
        the session signing key comes from env/vault and fails closed when absent
        in production.
\end{itemize}

\subsection{Audit integrity}
The \code{audit\_log} table (\code{db/schema.sql}) is \textbf{append-only and
hash-chained}: a \code{BEFORE INSERT} trigger computes \code{prev\_hash} /
\code{row\_hash}, and \code{UPDATE} / \code{DELETE} are blocked by trigger. CI
actively smoke-tests that an \code{UPDATE} is rejected. Every mutation and every
scope denial is recorded with the \emph{server-derived} actor --- the trail is
tamper-evident.

\subsection{Agent-layer enforcement (the Claude Code plugin)}
The \code{plugin/} package lets Claude drive the console while preserving the
boundary on the server:
\begin{itemize}
  \item \textbf{MCP server} (\code{plugin/mcp/vantage\_server.py}): a thin
        forwarder over the API; there is \emph{no} exploit/attack tool (the smoke
        test asserts no offensive verb leaked), and \code{request\_scan} forwards
        to the fail-closed server scope gate.
  \item \textbf{Scope-guard hook} (\code{plugin/hooks/scope\_guard.py}): a
        \code{PreToolUse} hook that parses (never executes) a Bash command,
        detects scanner tools, extracts target hosts, and \textbf{denies}
        (fail-closed) any scan against a host not in the allowlist
        (\code{approved\_scope.txt} $\cup$ live \code{GET /api/assets}); external
        \code{WebFetch} becomes ``ask''.
\end{itemize}

\section{Control-to-verification traceability}
Every control is backed by a test that runs in CI.

\begin{longtable}{L{4.6cm} L{6.0cm} L{3.5cm}}
\toprule
\textbf{Control} & \textbf{Verification} & \textbf{Result} \\
\midrule
\endhead
Scope gate (fail-closed) & \code{test\_api\_auth\_integration.py},
\code{test\_scan\_diff.py} & 403 out\_of\_scope \\
\addlinespace
No exploit phase/verb & \code{test\_scope\_invariants.py} + CI grep guard & green \\
\addlinespace
OIDC \code{id\_token} validation & \code{api/test\_auth.py} (mock JWKS) & alg-pinned, claims checked \\
\addlinespace
Session forgery / fail-closed & \code{api/test\_auth.py} & raises w/o secret \\
\addlinespace
RBAC matrix & \code{test\_fp\_exception\_workflow.py}, \code{test\_escalations\_api.py} & viewer 403 / role 200 \\
\addlinespace
Server-derived actor & \code{test\_api\_auth\_integration.py} & audit = session actor \\
\addlinespace
Report owner-scoping & \code{test\_api\_auth\_integration.py} & non-owner 403 \\
\addlinespace
LDAP injection guard & \code{api/test\_auth.py} & metachar email rejected \\
\addlinespace
Open-redirect guard & \code{api/test\_auth.py} & backslash bypass blocked \\
\addlinespace
XXE-safe parsing & adapter tests + \code{test\_scope\_invariants.py} & defusedxml only \\
\addlinespace
Audit append-only / hash-chain & schema smoke test (CI) & UPDATE rejected \\
\addlinespace
Dual-password PDF & \code{reporting/test\_reporting.py} & AES-256 R6 \\
\addlinespace
Scope-guard hook & \code{plugin/hooks/test\_scope\_guard.py} & 9/9 (incl. fail-closed) \\
\addlinespace
MCP boundary (no exploit tool) & \code{plugin/mcp/test\_server\_smoke.py} & no offensive verb \\
\bottomrule
\end{longtable}

\section{Security review methodology and results}
Changes are reviewed by an automated, three-phase security review:
\begin{enumerate}
  \item an identification pass traces untrusted input (request bodies/params,
        OIDC claims, LDAP responses, scanner output, report parameters) to
        sensitive sinks (SQL, file paths, the token/JWKS validation, HTML);
  \item parallel false-positive filters re-examine each candidate;
  \item findings below confidence 8/10 are dropped.
\end{enumerate}
The categories examined are input validation (SQLi, command injection, XXE, path
traversal, SSRF), authentication/authorization, cryptography/secrets, code
execution/deserialization, and data exposure. Noise classes (DoS, secrets-at-rest,
rate-limiting, hardening-without-exploit, dependency CVEs) are excluded.

\paragraph{Outcome.} Across reviews of the full server-side surface and the plugin
code, the codebase is currently \textbf{clean} at the reporting bar. The only
finding to date --- a low-severity open redirect via a backslash bypass in
\code{\_safe\_next} --- was remediated (rejecting backslash/CR/LF and re-validating
the parsed network location) and is covered by a regression test.

\appendix
\section{Build and run}
\begin{itemize}
  \item Stack: \code{docker compose up --build} --- console at
        \href{http://localhost:8137}{localhost:8137}, API at \code{:8138},
        Postgres at \code{:5432} (schema auto-applied).
  \item Tests: each control's test is listed in \S5 and runs in GitHub Actions CI.
\end{itemize}

\section{Production configuration (env)}
\begin{itemize}
  \item \code{AUTH\_REQUIRED=true} --- enforce real OIDC (default off in the
        reference build).
  \item \code{OIDC\_ISSUER}, \code{OIDC\_CLIENT\_ID}, \code{OIDC\_CLIENT\_SECRET},
        \code{OIDC\_REDIRECT\_URI} --- IdP configuration.
  \item \code{SESSION\_SECRET} (required when auth is on), \code{SESSION\_COOKIE\_SECURE=true}.
  \item \code{VANTAGE\_GROUP\_ROLE\_MAP} --- AD-group to role mapping;
        \code{LDAP\_*} for \code{memberOf} resolution.
  \item \code{DATABASE\_URL} --- Postgres for the persistence overlay and the
        hash-chained audit log.
\end{itemize}

\vfill
\noindent\rule{\textwidth}{0.4pt}\\
\small Generated as control evidence for the Vantage reference build. Source of
truth: the repository tests and \code{docs/auth-contract.md} / \code{docs/api-contract.md}.
\end{document}
